---
title: "Lab 06 - Advanced Graphics"
author: "Francisco Santamarina"
date: "October 25, 2016"
output: html_document
---
Load the necessary packages and dataset.
```{r}
##load data
library( Lahman )
library( dplyr )
data( Teams )
#ave.so <- Teams$SO / Teams$G
years <- filter(Teams, Teams$yearID >= 1900 & Teams$yearID <=2012 & Teams$yearID != "NA")
ave.so <- years$SO / years$G
```

#Part I

## Question 1
**Plot average strike-outs by year. Use the gray() function to select a color, and add transparency.**<br>

```{r}
plot( x=years$yearID, y=ave.so,
      col=gray(0.5,0.3) #shade of gray (0=black, 1=white) and opacity
    )
```

## Question 2
**Use appropriate x and y axis conventions and labels.**<br>
```{r}
plot( x=years$yearID, y=ave.so,
      col=gray(0.85, alpha=0.8 ), #shade of gray (0=black, 1=white) and opacity
      xlim=c(1900,2012), 
      ylim=c(0,10), 
      yaxt='n',
      xlab="",
      ylab="",
      #main="Hits and Wins per Seaso, 1990 to Present",
      #xlab="Hits by Team Per Season", #x-ax
      
      abline( h=seq(0,9,1), #vertical line written ([lowest value], [highest value], [intervals to draw line])
              lty=3, #line type
              col=gray(0.5,1) #color of line
              ),
      #col="green4", #color
      pch=16, #indicates the plot symbol to use
      cex=1, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
axis( side=4, at=1:9, labels = c(1:9), las=1, col.axis=gray(0.85), tick = FALSE )
```

## Question 3
**Calculate the league average using tapply() or group_by() year, and add it to the graph. Use the argument type="b" to connect the dots.**<br>

```{r}
plot( x=years$yearID, y=ave.so,
      col=gray(0.85, alpha=0.8 ), #shade of gray (0=black, 1=white) and opacity
      xlim=c(1900,2012), 
      ylim=c(0,10), 
      yaxt='n',
      xlab="",
      ylab="",
      #main="Hits and Wins per Seaso, 1990 to Present",
      #xlab="Hits by Team Per Season", #x-ax
      
      abline( h=seq(0,9,1), #vertical line written ([lowest value], [highest value], [intervals to draw line])
              lty=3, #line type
              col=gray(0.5,1) #color of line
              ),
      #col="green4", #color
      pch=16, #indicates the plot symbol to use
      cex=1, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
axis( side=4, at=1:9, labels = c(1:9), las=1, col.axis=gray(0.85), tick = FALSE )

league.ave <- tapply( ave.so, years$yearID, mean, na.rm=T )
#points( league.ave, type="b", col="blue3", pch=18)
points( names(league.ave), league.ave, type="b", col="deepskyblue4", pch=16)
```

## Question 4
**Reproduce at least two of the narrative texts ("US enters World War I", etc.). Note that a line break within text is created by including "\n" in your string.**<br>

```{r}
plot( x=years$yearID, y=ave.so,
      col=gray(0.85, alpha=0.8 ), #shade of gray (0=black, 1=white) and opacity
      xlim=c(1900,2012), 
      ylim=c(0,10), 
      axes=F,
      xlab="",
      ylab="",
      #main="Hits and Wins per Seaso, 1990 to Present",
      #xlab="Hits by Team Per Season", #x-ax
      
      abline( h=seq(0,9,1), #vertical line written ([lowest value], [highest value], [intervals to draw line])
              lty=3, #line type
              col=gray(0.7,1) #color of line
              ),
      #col="green4", #color
      pch=16, #indicates the plot symbol to use
      cex=1, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
axis( side=4, at=1:9, labels = c(1:9), las=1, col.axis=gray(0.5), tick = FALSE )
axis( side=1, at=seq( 1900, 2012, 10 ),  las=1, col.axis=gray(0), tick = TRUE )


league.ave <- tapply( ave.so, years$yearID, mean, na.rm=T )
#points( league.ave, type="b", col="blue3", pch=18)
points( names(league.ave), league.ave, type="o", col="dodgerblue3", pch=16)

text( x= 1963, y=2.5, labels="Strike zone enlarged\nfrom 1963-68.", col=gray(0.5) )
text( x= 1917, y=1, labels="U.S. enters\nWorld War I.", col=gray(0.5) )

lines( x=1963,  y=5.805269, col=gray(0.5) )

```

#Part II

## Question 1
**Using R Shiny tools, add an input panel that allows users to select a team. Once selected, add the team trendline to the graph.**<br>

```{r}
selectInput( 
             inputId='team', 
             label='Choose a Team', 
             choices=choices = unique( years$name )
           )
#input.ave <- filter(  )
#points( names(league.ave), league.ave, type="b", col="dodgerblue3", pch=16)

```

## Question 2
**Report the selected team name on the graph.**<br>

```{r}
plot( x=postteams$ERA, y=postteams$W,
      xlim=c(2,8), #limits the range of data for variable x
      ylim=c(20,140), #limits the range of data for variable x
      main="ERA and Wins Per Season, 1990 to Present", #title
      xlab="Earned Runs Average Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="darkviolet", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
text( x= 6.5, y=53.4, labels= "Outlier 1", pos=4, col="gray0" )
text( x= 3.6, y=116.8, labels="Outlier 2", pos=4, col="gray0" )

lines( lowess(postteams$ERA, postteams$W), col="chocolate4", lwd=1)
```
