---
title: "Lab 04 - Describing Data"
author: "Francisco Santamarina"
date: "September 27, 2016"
output: pdf_document
---
Load the necessary packages and dataset.
```{r}
##load data
library(dplyr)
library(pander)
dat <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
```

# Question 1
**What class is the dataset?**
"data.frame"
```{r}
class(dat)
```

**How many rows of data are there?**
41502 rows of unique data
```{r}
nrow(dat)
```

**How many variables?**
64 variables
```{r}
length(dat)
```

**How would you check the class of a variables in the dataset?**
```{r}
names(dat)
class(dat$TAX_ID)
```

# Question 2
**2.	Convert your dataset into a tibble using the tbl_df() function. What is the class of the dataset now? How can you check rows, columns, and class of variables for the tibble?**<br>
```{r}
tbl.dat <- tbl_df(dat)
```

**What is the class of the dataset now?**<br>
It is a tbl and inherits the class data.frame
```{r}
class(tbl.dat)
typeof(tbl.dat)
```

**How can you check rows, columns, and class of variables for the tibble?**<br>
```{r}
#Rows/Observations, Columns/Variables, and class:
glimpse(tbl.dat)
```

# Question 3
**3.	Create a subset of the data by retaining the following set of variables:**<br>

* Acres, LandUse, AmtDelinqu,  AssessedLa, VacantBuil, YearBuilt, Nhood<br>

```{r}
littletbl <- select(.data=tbl.dat, Acres, LandUse, AmtDelinqu, AssessedLa, VacantBuil, YearBuilt, Nhood)
```

# Question 4
**4.	Drop the cases there the year of construction (YearBuilt) is reported as zero using the filter() function.**<br>

```{r}
no.0.littletbl <- filter(.data = littletbl, YearBuilt != 0)
```

# Question 5
**5.	Create a new variable that measures the assessed land value (AssessedLa) per acre.**<br>

```{r}
alv.acre <- mutate(.data = no.0.littletbl, AssessedLa.acre = AssessedLa / Acres)
```

# Question 6
**6.	Which neighborhood has the highest average land value per acre? Use the group_by() and summarise() functions to answer this question.**<br>
Downtown, at an average of $895,080.30 per acre
```{r}
#g.alv.acre <- group_by(alv.acre, Nhood)
#summ.g <- summarise( g.alv.acre, AvgALV.acre = mean( AssessedLa.acre))
#arrange(summ.g, desc( AvgALV.acre) )

arrange(summarise( group_by(alv.acre, Nhood), AvgALV.acre = mean( AssessedLa.acre)), desc( AvgALV.acre) )
```

# Question 7
**7.	OPTIONAL: Create a new variable that splits the year of construction up by decade using the cut() function. Which decade produced single family homes with the highest assessed value (AssessedVa)?**<br>

```{r}
```
