---
title: "Lab 03 - Merging Data"
author: "Francisco Santamarina"
date: "September 10, 2016"
output: pdf_document
---
Load the necessary packages and dataset.
```{r}
##load data
library(Lahman)
```

# Question 1
**What is the average salary of a player on the Oakland A's in 2002?**

$1,482,000

**A player on the Boston Red Sox that same year?**

$3,612,000

**What is the total budget of each franchise in 2002?**
Oakland A's: $40,004,167
Boston Red Sox: $108,366,060

To answer the question, merge the Salaries and Teams datasets.

```{r}
data( Master )
data( Teams )
data( Salaries )
dat.salaries.teams <- merge( Salaries, Teams )
OAK2002.salaries <- dat.salaries.teams$name == "Oakland Athletics" & dat.salaries.teams$yearID == "2002"
#mean( dat.salaries.teams$salary [OAK2002.salaries], na.rm=T )#produces weird values
summary(dat.salaries.teams$salary [OAK2002.salaries])

#find players on Red Sox in 2002
#merge that with salary
#find average

BOS2002.salaries <- dat.salaries.teams$name == "Boston Red Sox" & dat.salaries.teams$yearID == "2002"
#mean( dat.salaries.teams$salary [BOS2002.salaries], na.rm=T )#produces weird values
summary(dat.salaries.teams$salary [BOS2002.salaries])


#find the franchises in 2002
#find the salaries for each one
#sum the salaries for each one
#print as a table

sum( dat.salaries.teams$salary [OAK2002.salaries] )
sum( dat.salaries.teams$salary [BOS2002.salaries] )

```

# Question 2
**2.	Which variables are used for the default merge? Why would the following merge be problematic?**<br>

merge(  Teams, Salaries, by.x="teamID", by.y="teamID" )

The variables being used are teamID from Teams and teamID from Salaries. 
* This uses Teams as the primary table. Any independent players whose salaries are not associated to a team, or with a teamID, will not appear in the merged dataset. 
* This assumes that teamID are the same uniqueID and are the appropriate variables to use
* This merge will not include teams without salaries.

# Question 3
**3.	Which city (birth city) has produced the most baseball hall-of-famers? Use the Master and HallOfFame datasets.**<br>

San Francisco, with 128 players

To check which variables will be used for the default merge:
intersect( names(HallOfFame), names(Master) )
```{r}
library(Lahman)
data( Master )
data( HallOfFame )
intersect( names(HallOfFame), names(Master) )
ls(Master)
master.halloffame <- merge( Master, HallOfFame, by.x = "playerID", by.y = "playerID")
sort( table( master.halloffame$birthCity))

```
