---
title: "Lab 05 - Describing Data"
author: "Francisco Santamarina"
date: "October 13, 2016"
output: pdf_document
---
Load the necessary packages and dataset.
```{r}
##load data
library(dplyr)
library(pander)
library("Lahman")
data("Teams")
```

# Question 1
**1.	Drop all of the data before 1990.**<br>

```{r}
postteams <- filter(Teams, yearID>=1990)
summary(postteams)
```

# Question 2
**2.	Examine the relationship between Hits and Wins using a scatterplot. Use meaningful labels and a visually appealing style.**<br>
```{r}
plot( x=postteams$H, y=postteams$W,
      xlim=c(750,2000), #limits the range of data for variable x
      main="Hits and Wins per Seaso, 1990 to Present", #title
      xlab="Hits by Team Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="green4", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )

```

# Question 3
**3.	Annotate two outliers on the graph with the teamID. You can identify the points using the identify function and clicking on the graph near the points to return their positions: identify( x, y )**<br>

```{r}

plot( x=postteams$H, y=postteams$W,
      xlim=c(0000,2000), #limits the range of data for variable x
      main="Hits and Wins per Season, 1990 to Present", #title
      xlab="Hits by Team Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="green4", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )

text( x= 1306, y=43, labels= "Outlier 1", pos=4, col="gray0" )
text( x= 1630, y=114, labels="Outlier 2", pos=4, col="gray0" )
```

# Question 4
**4.	Examine the relationship between ERA and Wins using a scatterplot. Use meaningful labels and a visually appealing style.**<br>

```{r}
plot( x=postteams$ERA, y=postteams$W,
      xlim=c(2,8), #limits the range of data for variable x
      ylim=c(20,140), #limits the range of data for variable x
      main="ERA and Wins Per Season, 1990 to Present", #title
      xlab="Earned Runs Average Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="darkviolet", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
```

# Question 5
**5.	Annotate two outliers on the graph with the teamID.**<br>

```{r}

identify( 8, 0 )
identify( 7, 55 ) 

plot( x=postteams$ERA, y=postteams$W,
      xlim=c(2,8), #limits the range of data for variable x
      ylim=c(20,140), #limits the range of data for variable x
      main="ERA and Wins Per Season", #title
      xlab="Earned Runs Average Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="darkviolet", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
text( x= 6.5, y=53.4, labels= "Outlier 1", pos=4, col="gray0" )
text( x= 3.6, y=116.8, labels="Outlier 2", pos=4, col="gray0" )
```

# Question 6
**6.	BONUS - add a trend line to the scatterplot to highlight the relationship.**<br>

```{r}
plot( x=postteams$ERA, y=postteams$W,
      xlim=c(2,8), #limits the range of data for variable x
      ylim=c(20,140), #limits the range of data for variable x
      main="ERA and Wins Per Season, 1990 to Present", #title
      xlab="Earned Runs Average Per Season", #x-axis label
      ylab="Wins Per Season", #y-axis label
      col="darkviolet", #color
      pch=20, #indicates the plot symbol to use
      cex=.8, #determines size of the symbols/plots
      bty="n" #gets rid of the box around the plot
    )
text( x= 6.5, y=53.4, labels= "Outlier 1", pos=4, col="gray0" )
text( x= 3.6, y=116.8, labels="Outlier 2", pos=4, col="gray0" )

lines( lowess(postteams$ERA, postteams$W), col="chocolate4", lwd=1)
```
