---
title: "Lab 01 - Load a Package"
author: "Francisco Santamarina"
date: "September 5, 2016"
output: html_document
---
The following code was written assuming that the user has already installed the Lahman package, but it has not been opened.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Question 1
What is the average weight of an MLB player?

**Answer:** Mean = 186
```{r}
library("Lahman")
data("Master")
summary(Master$weight)
```
***
### Question 2
How many MLB players have come from the Netherlands?

**Answer:** 12 MLB players have come from the Netherlands
```{r}
library("Lahman")
data("Master")
table(Master$birthCountry)
```
***
### Question 3
How tall was the tallest player in MLB? The shortest?

**Answer:** The tallest player in MLB was 83, and the shortest was 43
```{r}
library("Lahman")
data("Master")
summary(Master$height)
```
***
### Question 4
What percentage of players switch hit (bat from both sides of the plate)?

**Answer:** 6.59% of players are switch hitters. 
```{r}
library("Lahman")
data("Master")
summary(Master$bats)
switchhitter <- 1163
total.hitter <- (1163+4879+11614)
percent.switchhitter <- (switchhitter)/(total.hitter)
paste((round((percent.switchhitter*100),digits = 2)),"%")
```
***
### Question 5
**OPTIONAL:** How many players are right-handed when they throw, but left-handed when they bat?
```{r}
##library("Lahman")
##data("Master")
## identify the list of players that are right-handed throwers. Set that value to "rightthrowers"
## identify the list of players that are left-handed batters. Set that value to "leftbatters"
## compare the list of values, and produce a summary of the count of combinations, such as left batter & left pitcher, right batter & right pitcher, etc. 
```
