---
title: "Lab 02 - Logical Operators_FJS Comments"
author: "Francisco Santamarina"
date: "September 9, 2016"
output: html_document
---
Load the necessary packages and dataset.
```{r}
##load data
library(dplyr)
library(pander)
dat <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
```

# Question 1
What is the total assessed value of all of the taxable land in Syracuse?<br>
* $7,538,424,950
```{r}
totalassessedva <- sum(dat$AssessedVa, na.rm=T)
as.numeric((totalassessedva))
```

# Question 2
What proportion of land parcels is not taxable? What is the value of these non-taxable parcels? <br>
1. 12.28%
2. $3,022,568,602
```{r}
# Proportion of land parcels that are not taxable
unique( as.character( dat$LandUse ))
#taxable <- dat$LandUse == c( "Single Family", "Commercial", "Two Family", "Three Family", "Apartment", "Industrial" )
#     taxable <- dat$LandUse %in% c( "Single Family", "Commercial", "Two Family", "Three Family", "Apartment", "Industrial" )
taxable <- dat$LandUse=="Single Family" | dat$LandUse=="Two Family" | dat$LandUse=="Three Family" | dat$LandUse=="Commercial" | dat$LandUse=="Apartment" | dat$LandUse=="Industrial"
percent.nottaxable <- sum( !taxable ) / length(dat$LandUse)
percent.nottaxable
paste0((round((percent.nottaxable*100),digits = 2)),"%")

#chris code
#chris.code <- dat$LandUse=="Single Family" | dat$LandUse=="Two Family" | dat$LandUse=="Three Family" | dat$LandUse=="Commercial" | dat$LandUse=="Apartment" | dat$LandUse=="Industrial"
#total.chris <- sum(!chris.code)
#total <- total.chris/nrow(dat)
#total

#Value of the non-taxable parcels
sum (dat$AssessedVa [!taxable])

#  value.nontax <- !taxable & dat$AssessedVa
#  sum( !taxable)

#  summary( not.taxable, dat$AssessedVa)
```

# Question 3
Which property has the highest delinquient tax bill?
* Property with row location 22469 in the data set has the highest delinquent tax bill.
```{r}
which.max(dat$AmtDelinqu)
max.delinquent <- dat[22469,]
pander( max.delinquent )
```

# Question 4
Which of these neighborhoods listed below has the highest proportion of vacant buildings? Use the Nhood
and VacantBuil variables.
* Strathmore, at 8.76% of vacancies.
```{r}
# empty.wcott <- dat$Nhood == "Westcott" & dat$VacantBuil == "Y"
# proportion.wcott <- sum(empty.wcott, na.rm=T) / sum(dat$Nhood == "Westcott", na.rm=T)
# proportion.wcott
# paste0((round((proportion.wcott*100),digits = 2)),"%")

find.vacantbld <- function(mycrib)
{
  empty.nhood <- dat$Nhood == mycrib & dat$VacantBuil == "Y"  
  prop.emptynhood <- sum(empty.nhood, na.rm=T) / sum(dat$Nhood == mycrib, na.rm=T)
  return(prop.emptynhood)
}

emptywcott <- find.vacantbld("Westcott")
emptynside <- find.vacantbld("Northside")
emptysmore <- find.vacantbld("Strathmore")
emptysside <- find.vacantbld("Southside")
empty.4hoods <- c(emptywcott,emptynside,emptysside, emptysmore)

names(empty.4hoods) <- c("Westcott", "Northside", "Strathmore", "Southside")
which.max(empty.4hoods)

index<- which.max(empty.4hoods)
prop.mostempty <- empty.4hoods[index]
paste0((round((prop.mostempty*100),digits = 2)),"%")
```

# Question 5
How many single family homes worth more than $100,000 each are located in each of the four neighborhoods
above? Use the AssessedVa variable?
* Westcott has 208 homes worth more than $100K
* Northside has 5 homes worth more than $100K
* Strathmore has 461 homes worth more than $100K
* Southside has 0 homes worth more than $100K
```{r}
find.100khomes <- function(mycrib){
  singlehome.inhood <- dat$LandUse=="Single Family" & dat$Nhood == mycrib & dat$AssessedVa > 100000 
  return(sum(singlehome.inhood))
}
find.100khomes("Westcott")
find.100khomes("Northside")
find.100khomes("Strathmore")
find.100khomes("Southside")

```

