---
title: 'Generating geoJSON file of Cortland City'
author: "Francisco Santamarina"
date: "June 09, 2017"
output:
  html_document:
    code_fold: hide
    df_print: paged
    keep_md: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )
```

## Load the packages
```{r}
# Libraries contain functions not in the base version of R allowing you greater flexibility and customization
# These specific libraries are recommended by Professor Lecy for doing regressions.

library( ggplot2 ) # Cool plots
library( knitr ) # for formatting tables
library( magrittr ) # cool ways of coding things that does more with less. Check it out!
library( shiny )
```

## Set the working directory
```{r, eval=FALSE}
# Set the working directory
getwd()
setwd("~/Graduate School/Capstone/Census")

```


## Load the necessary datasets
```{r, results='hide'}
# Read the .csv of the tracts data file into your R environment
tracts <- read.csv( "CensusTracts.csv", header = T )

# Read the .csv of the variables data file into your R environment
variables <- read.csv( "CensusVariables.csv", header = T )

# Read the .csv of the ACS data file into your R environment
dat <- read.csv( "ACSData.csv", header = T )

var_names <- as.data.frame( unique( names(dat) ) )
var_names <- as.data.frame( var_names[ -c(1:4,93:94), ] )
years <- unique( dat$YEAR )
```

## Clean the Census Tracts data
```{r}
# CLEANING THE TRACTS DATA
# Return the structure of the vector "dat"
#str( tracts )

tracts_county <- tracts[ -13, ]
tracts_county <- as.data.frame( tracts_county )
str( tracts_county )

tracts_city <- tracts[14:18,]
tracts_city <- as.character( factor( tracts_city ) )
#tracts_city <- as.data.frame( tracts_city )
#str( tracts_city )

rm( tracts )
```

# Prep the Maps
Shapefile zip located here: https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2010&layergroup=Census+Tracts
**_Download the 2010 version for compatibility with the 2010 census data_**

## Map libraries
```{r}
library( geojsonio ) # for manipulating GeoJSON files
library( maptools ) # for manipulating shapefile data, can also use to import shapefiles
library( raster ) # another package for importing
library( rgdal ) # for importing shapefiles into R
library( spdplyr ) # for manipulating attribute data within shapefiles, can use DPLYR functions on shapefiles
library( rmapshaper ) # for manipulating geometries within GeoJSON files
library( sp ) # for manipulating spatial points data
library( tmaptools ) # another way to manipulate/read in shapefiles
```

## Clean the shapefiles
Set the working directory
```{r}
#setwd("~/Graduate School/Capstone/Census/Cortland_County_tl_2016_36023_addrfeat")
```
Load the county geoJSON file
```{r}
# Download the code:
  script <- getURL("https://raw.githubusercontent.com/fjsantam/Maxwell-R-Labs/master/MPA_Workshop/cortland.geojson", ssl.verifypeer = FALSE)
  # Read the code as a shapefile:
  county <- readOGR( script )
  # Remove the holding variable "script":
  rm( script )

#Confirm you have a shapefile:
class( county )

#See how it looks: 
plot( county )

#Check out the data in your vector using "@" to call the data within the shapefile:
head( county@data )

#proj4string( county ) <- CRS("+proj=longlat +datum=WGS84")

#Alternate code: 
county = read_shape("C:/Users/franc/Documents/Graduate School/Capstone/Census/Cortland_County_tl_2010_36023_tract10/tl_2010_36023_tract10.shp")

#Plot the shapefile:
qtm(county)
```

Create a subset of the county map that only shows the city
```{r}
city <- county[ county@data$GEOID10 == tracts_city, ]
city = county[ county$GEOID10 == "36023970500", ]

city <- subset( county, GEOID10 %in% tracts_city )
plot(city)
```


V1: Convert Shapefile to GeoJSON
We will be following the steps here: https://blog.exploratory.io/creating-geojson-out-of-shapefile-in-r-40bc0005857d
```{r}
city <- city %>% rename( city_id = GEOID10 )

# Convert to GeoJSON
city <- geojson_json(city)

# Confirm the class
class( city )

# See how it looks
plot(city)

geojson_write( city, file = "~/cortland_city.geojson" )
```

V2: Convert Shapefile to GeoJSON **EASIER VERSION**
We will be following the steps here: https://github.com/Robinlovelace/Creating-maps-in-R/blob/master/R/shp-to-geojson.R
```{r}
geojson_write( city, file = "C:/Users/franc/Documents/Graduate School/Capstone/Census/cortland.city.geojson" )

# Should return: "Success! File is at (directory specified above)"
```

## Test GeoJSON file
```{r}
cortland.city <- readOGR( dsn = "C:/Users/franc/Documents/Graduate School/Capstone/Census/cortland.city.geojson" )
plot(cortland.city)
```
